---
title: "Replication timing"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Replication timing

```{r, results="hide", message=FALSE, warning=FALSE, fig.show="hide", echo=FALSE}
library(plotly)
library(RSQLite)
```

Let's see how this new dataset looks:

```{r, message=FALSE, warning=FALSE}
conn <- dbConnect(RSQLite::SQLite(), "../../Results/yeast.sqlite")
replicationTiming <- dbGetQuery(conn, "SELECT l.Gene, x, y, z, ReplicationTiming FROM Loci l JOIN ReplicationTiming r ON l.Gene = r.Gene")
plot_ly(x=replicationTiming$x, y=replicationTiming$y, z=replicationTiming$z, type="scatter3d", mode="markers", text=replicationTiming$Gene, color=replicationTiming$ReplicationTiming)
dbDisconnect(conn)
```

## Metric used

We use standard deviation of replication timing in a sphere as a metric.

This signal is so smooth that almost all possible spheres have a far below metric than the randomized set of same size.

Maybe it makes more sense to find the rare spheres where smoothness is disrupted:

```{r, message=FALSE, warning=FALSE}
sql <- "SELECT l.*, CASE WHEN Field IS NULL THEN 'None' ELSE Field END AS Field FROM Loci l LEFT JOIN ReplicationTimingFields f ON l.Gene = f.Gene"
fieldColors <- c("#FF0000", "#00FF00", "#0000FF", "#FFFF00", "#00FFFF", "#4488AA", "#FF00FF", "#000000")
conn <- dbConnect(RSQLite::SQLite(), "../../Results/yeast.sqlite")
replicationTimingFields <- dbGetQuery(conn, sql)
plot_ly(x=replicationTimingFields$x, y=replicationTimingFields$y, z=replicationTimingFields$z, type="scatter3d", mode="markers", text=replicationTimingFields$Gene, color=replicationTimingFields$Field, colors=fieldColors)
dbDisconnect(conn)
```

## Overlaps

* ReplicationTimingFields:C has maximum overlap 100% with MotifFields:A (one is strict subset of the other).
* ReplicationTimingFields:B is 100% within Tethys.

## Worries

Field C is the same as MotifFields A, which was again the weird one, having more variance in TFs than expected by chance.
These 123 genes are all located in chrXII, from ~300Kb to ~800KB, with a 200Kb gap between ~500-700.

Maybe we have the wrong (index space) coordinates for these genes? That would explain having significantly varied signals for both signal types.

Alternatively, this weird behavior is because of the rDNA that is present there.


```{r, message=FALSE, warning=FALSE}
sql <- "SELECT Gene, Chromosome, Start,End FROM Loci WHERE Gene IN (SELECT Gene FROM ReplicationTimingFields WHERE Field = 'C') ORDER BY Start"
conn <- dbConnect(RSQLite::SQLite(), "../../Results/yeast.sqlite")
weirdGenes <- dbGetQuery(conn, sql)
dbDisconnect(conn)
weirdGenes
```




